---
title: "Light transmission"
author: "Mike Paterson"
date: "30/11/2020"
output: 
  html_document:
    number_sections: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(here)
library(gt)
library(broom)
```
# Data files

F_par.csv file w light profiles from ELA lakes (1986-2019)
attenuation.csv old_kd_estimates for ELA lakes (1968-1985)
F_secchi.csv secchi disc measures from ELA lakes (1968-2019)

```{r, message=FALSE}
light_profiles <- read_csv(here("data_wrangling/data", "F_par.csv"),guess_max = 10000)
light_profiles$date <- as_date(light_profiles$date,"Ymd")

# WARNING: the F_PAR file does not have sub-basins in it, so needs to be updated for lakes with 2 or more basins

old_kd_estimates <- read_csv(here("data_wrangling/data", "attenuation_a.csv"), guess_max = 10000)
old_kd_estimates$date <- as_date(old_kd_estimates$date,"Ymd")
secchi <- read_csv(here("data_wrangling/data", "field_observations.csv"), guess_max = 10000)
secchi$date <- as_date(secchi$date,"Ymd")
secchi$lake <- as.character(secchi$lake)
secchi <- secchi %>%
  select(c(lake, sublocation, station, date, year, start_time, epi_depth, planar_thermocline_depth, secchi, integrated_epi_start_d, integrated_epi_end_d))
```
Have a look at the data from one year from one lake.  If you want you could use PURR to generate plots for all years and lakes to scan through.  I generally look at the data and look for anomalous values and then check them (and remove them, if necessary)

```{r, message=FALSE, warning=FALSE}
light_profiles %>%
  mutate(ln_light_ratio = log(light_percent/100)) %>%
  filter(lake == 239, year == 2001, depth > 0) %>% # select L239 and 1988 as an example to look at the data
  ggplot() +
  facet_wrap(~date) +
  geom_point(aes(x = depth, y = ln_light_ratio)) +
  geom_smooth(aes(x = depth, y = ln_light_ratio), method = "lm")
```
# Fit linear models 

log(percent light) vs depth for all dates, Lake 239

```{r, message=FALSE, warning=FALSE}
light_profiles_count <- light_profiles %>%
    group_by(lake, date) %>%
    summarize(count = n()) %>%
    ungroup()

light_profiles <- left_join(light_profiles, light_profiles_count)

light_extinction_models <- light_profiles %>%
  mutate(year = year(date)) %>%
  filter(count>3) %>%   # only include profiles with at least 3 measurements
#  filter(year > 2013 & year < 2020) %>%
  mutate(ln_light_ratio = log(light_percent/100)) %>%
  filter(flag != "yes", depth > 0) %>%     # these estimates don't use the air values 
  group_by(lake, date) %>% 
  nest() %>% 
  mutate(model = pmap(list(data),
                      ~ lm(ln_light_ratio ~ depth, data = ..1 )
                      )
         )
```
# Extract slope (Kd), intercept, r2, and p value for all regressions

```{r, message=FALSE, warning=FALSE}
kd_results <- light_extinction_models %>% 
  mutate(slope = map_dbl(.x = model,
                             ~ round(tidy(.x)$estimate[2], digits = 3)),
         intercept = map_dbl(.x = model,
                         ~ round(tidy(.x)$estimate[1], digits = 3)),
         r_squared = map_dbl(.x = model,
                             ~ round(glance(.x)$r.squared, digits = 3)),
         p_value = map_dbl(.x = model,
                             ~ round(glance(.x)$p.value, digits = 5))
         
  )

print(kd_results)
```
# Extract Kd values to a csv file

```{r, message=FALSE, warning=FALSE}
kd_results$date <- as_date(kd_results$date,"Ymd")
kd_results <- kd_results %>%
  select(lake, date, slope, intercept, r_squared, p_value) %>%
   mutate(year = year(date)) %>%
  mutate(month = month(date)) %>%
  write_csv(here("data_wrangling/wrangled_data", "kd-results_all_lakes.csv"))
```

# Join new estimates with old estimates to generate full data set covering all years. Then join with secchi disc data
```{r, message=FALSE, warning=FALSE}
old_kd_estimates$lake <- as.character(old_kd_estimates$lake)
kd_results$date <- as_date(kd_results$date,"Ymd")
kd_results$lake <- as.character(kd_results$lake)
combined_light_data <- full_join(secchi, old_kd_estimates) %>% 
   full_join(kd_results) 
print(combined_light_data)
```

```{r, message=FALSE, warning=FALSE}

combined_light_data3 <- combined_light_data %>%
  mutate(kd_best_fit = replace_na(kd_best_fit, 0))%>% # remove NAs from old Kd estimates
  mutate(kd_best_fit = kd_best_fit * -1) %>% 
  mutate(kd_combined_data = if_else(kd_best_fit < 0, kd_best_fit, slope)) %>% 
  #use previous estimate, if available, otherwise use calculated Kd
  mutate(kd_best_fit = na_if(kd_best_fit, 0)) %>% 
  select(lake, sublocation, date, year, month, jd, epi_depth, secchi, sublocation, station, kd_best_fit, kd_all, kd_combined_data)

combined_light_data3$date <- as_date(combined_light_data3$date,"Ymd")

combined_light_data3 %>%
write_csv(here("data_wrangling/wrangled_data", "combined_light_data_all_lakes.csv"))

seasonal_light <- combined_light_data3 %>% # calculate annual means by month for plotting
  select(lake, date, year, month, epi_depth, secchi, kd_combined_data) %>%
  pivot_longer(cols = c(-lake, -date, -year, -month), names_to = "parameter", values_to = "measure") %>%
  group_by(lake, year, month, parameter) %>%
  summarise(monthly_mean = mean(measure, na.rm = TRUE)) %>%
  mutate(month_name = case_when(
    month == "4" ~ "April",
    month == "5" ~ "May",
    month == "6" ~ "June",
    month == "7" ~ "July",
    month == "8" ~ "August",
    month == "9" ~ "September",
    month == "10" ~ "October",
    month == "11" ~ "November",
    month == "12" ~ "December"))

#Calculate monthly means (you may not want this)

seasonal_light %>%
   pivot_wider(names_from = parameter, values_from = monthly_mean) %>%
write_csv(here("data_wrangling/wrangled_data", "light_monthly_means_all_lakes.csv"))

#Calculate summer means (again, you may not want this)

annual_summer_light <- seasonal_light %>%
  mutate(summer_flag = if_else(month_name == "June" | month_name == "July" | month_name == "August", 1, 0)) %>%
  group_by(lake, year, parameter) %>%
  summarise(months = sum(summer_flag, na.rm = TRUE), summer_mean = mean(monthly_mean, na.rm = TRUE))

annual_summer_light$summer_mean[is.nan(annual_summer_light$summer_mean)]<-NA

annual_summer_light %>%
   pivot_wider(names_from = parameter, values_from = summer_mean) %>%
  filter(months == 3) %>% #include only if data from all 3 months
write_csv(here("data_wrangling/wrangled_data", "light_summer_means.csv"))
```

